---
title: 'Project 2: Class 607 Wine Quality'
author: "Blessing Anoroh"
date: "March 1, 2024"
output: html_document
---

This data is used to determine the physciochemicals of Wine Quality.

<font size=5><b>  Creating a .CSV file and Read data </font></b>

```{r}
library(readr)

wine <- read.csv("/Users/blessinga/Desktop/winequality-red.csv", header=TRUE, sep=",")
str(wine)

colnames(wine)
```


```{r}
# Checking for missing values
sum(is.na(wine))

# Summary statistics
summary(wine)
```

## Renaming column names

```{r}
library(tidyr)
library(dplyr)

wine2<-rename(wine,FxAc=fixed.acidity, VolAc=volatile.acidity, CtAc=citric.acid, ResiSug=residual.sugar, Chlo=chlorides, FrSulfDio= free.sulfur.dioxide, TtlSulfDio= total.sulfur.dioxide, Dens =density, pH= pH, Sulph= sulphates, Alc=alcohol, Qlty = quality)
head(wine2)

```

## Analysis of Wine Quality

<b> Analysis 1: Summary </b>

```{r}
#average values by wine quality
wine2_average_summary <- wine2 %>%
  group_by(Qlty) %>%
  summarise_all(.funs = list(mean))

# Display the summary
wine2_average_summary
```
<b>

Quality 7 has the highest average fixed acidity. \
Quality 3 has the highest average volatile acidity. \
Quality 3 and 4 have the lowest average citric acid. \
Quality 4 and 7 have the average higest Residual sugar. \
Quality 3 have the highest average Chloride. \
Quality 3 have the higest average free Sulfur dioxide. \
Quality 6 have the highest average total sulfur dioxide. \
The density for all qualities are quite similar as well as PH. \
Quality 3 have higher dense and pH and lower sulphates. \
Quality 8 have the highest alcohol level. \

ANALYSIS FROM THIS SUMMARY: \
Lower quality wines ratings 3 and 4 tend to have higher volatile acidity, lower citric acid, higher chlorides, higher density, higher pH, lower sulphates, and lower alcohol content. \
Higher quality wines ratings 7 and 8 exhibit the opposite characteristics, with lower volatile acidity, higher citric acid, lower chlorides, lower density, lower pH, higher sulphates, and higher alcohol content.

</b>

\
\
<b> Analysis 2: Bar Graph</b>

```{r}
library(ggplot2)

wine_sum3 <- wine %>%
  group_by(quality) %>%
  summarise(mean_fixed_acidity = mean(fixed.acidity),
            mean_volatile_acidity = mean(volatile.acidity),
            mean_citric_acid = mean(citric.acid),
            mean_residual_sugar = mean(residual.sugar),
            mean_chlorides = mean(chlorides),
            mean_free_sulfur_dioxide = mean(free.sulfur.dioxide),
            mean_total_sulfur_dioxide = mean(total.sulfur.dioxide),
            mean_density = mean(density),
            mean_pH = mean(pH),
            mean_sulphates = mean(sulphates),
            mean_alcohol = mean(alcohol))

wine_sum_long <- wine_sum3 %>%
  tidyr::pivot_longer(cols = -quality, names_to = "Property", values_to = "Mean_Value")

# Create bar plot
ggplot(wine_sum_long, aes(x = quality, y = Mean_Value, fill = Property)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Physicochemical Properties by Wine Quality",
       x = "Wine Quality",
       y = "Mean Value") +
  scale_fill_brewer(palette = "Set2") 
```

ANALYSIS: \
Focus on analysis Part 1 but this is an extra.\
 This analysis of the average values of physicochemical properties grouped by wine quality shows an in depth visual bar graph of the difference variables of high and low-quality red wines. This can help winemakers and customers know which wine choices they prefer to buy based on the variables.

\
\
