---
title: 'Project 2: Class 607  Property Sales'
author: "Blessing Anoroh"
date: "2024-03-04"
output: html_document
---

## Create CSV and read data

```{r setup}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)

property_sales <- read.csv("/Users/blessinga/Desktop/nyc-property-sales.csv", header=TRUE, sep=",")

str(property_sales)

summary(property_sales)

```

## Remove columns
```{r}
## Remove columns
nyc_sales <- property_sales %>%
  select(-c(BUILDING.CLASS.CATEGORY, TAX.CLASS.AT.PRESENT, BLOCK,LOT,EASE.MENT, BUILDING.CLASS.AT.PRESENT,ADDRESS, APARTMENT.NUMBER,ZIP.CODE,RESIDENTIAL.UNITS, COMMERCIAL.UNITS,TOTAL.UNITS,LAND.SQUARE.FEET,GROSS.SQUARE.FEET,TAX.CLASS.AT.TIME.OF.SALE,BUILDING.CLASS.AT.TIME.OF.SALE))

nyc_sales <- nyc_sales %>%
  na.omit()  

head(nyc_sales)

```


## Remame Columns and Analyze

<b> Analysis Part 1 </b>

```{r}
nyc_sales2 <-rename(nyc_sales, Borough = BOROUGH, Neighborhood = NEIGHBORHOOD, SalePrice = SALE.PRICE, SaleDate=SALE.DATE, YearBuilt = YEAR.BUILT) 


sales_summary <- nyc_sales2 %>%
  group_by(Borough) %>%
  summarise(mean_sale_price = mean(SalePrice),
            median_sale_price = median(SalePrice),
            min_sale_price = min(SalePrice),
            max_sale_price = max(SalePrice))

print(sales_summary)
```



<b> Analysis Part 2 </b>

```{r}
## line plot by borough it was meant to have colors by borough
library(ggplot2)
 ggplot(nyc_sales2, aes(x = SaleDate, y = SalePrice, color = Borough)) +
  geom_line() +
  labs(title = "Sale Price Trends Over Time",
       x = "Sale Date",
       y = "Sale Price") +
  theme_minimal()

```

## Analysis Conclusion 

<b>

Final Analysis based on Analysis Part 1 :

Borough 1 and 3 have the highest mean, median and max sales prices but most especially borough 1. Borough 1 most likely might be a borough where things are high in cost, from rent, food, groceries, and much more. It most likely might also be a borough near tourists areas and where many visitors tend to go to. Example Manhattan. Borough 1 properties have a higher value compared to the other boroughs.

</b>