---
title: "Week 3 Lab 607"
author: "Blessing Anoroh"
output: html_document
date: "02-08-2024"
---

## Question 1:
Problem: Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

<font size=3><b> Solution to Question 1 </font></b>

To solve Question 1 Problem; I first imported the raw college major data from Github, first using the link provided from fivethirtyeight website to open the article. <br/>
Then I loaded packages: dplyr , tidyverse and readr. Afterwards, I searched the majors data frame through filter. The filter uses regex to filter out majors that contain either ‘DATA’ or ‘STATISTICS’ and pulls it out into a new data frame called mathematical_majors.

```{r setup}
#import data and packages
library(tidyverse)
library(dplyr)
library(readr)
majors <- read.csv(url("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"))

#filter majors that contains "DATA" & "STATISTICS"
(mathematical_majors <- majors %>%
  filter(str_detect(Major, "DATA") | str_detect(Major, "STATISTICS")))


```


## Question 2

Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange".
 
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"  

[13] "olive"        "salal berry"  


Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")


<font size=3><b> Solution to Question 2 </font></b>


```{r }
fruits <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

fruit_vec <- paste0("\"", fruits, "\"")

fru_vec2 <- paste0(fruit_vec, collapse = ", ") 

fru_vec3 <- paste0("c(", fru_vec2, ")")

cat(fru_vec3)

```

 <font size=2><b> Steps </font></b>
 
So I had a hard time with this one a bit to make it look exactly like the way it is showed. I at first had it but minus the quotations around it. So I eventually came to this solution a different style. <br/>
I first created the string. The I had added backlash and quotation around each element (fruits), and defined it into a new vector called "fruit_vec". <br/>
Next, I collapsed the quoted words into a string, separating each word with a comma. Now identified as "fruit_vec2". <br/>
Then, I added the leading characters which is "c(.....)". <br/>
Finally, I parsed the backslashes as escape characters using cat function. The output is now in the desired format.



## Question 3

Describe, in words, what these expressions will match:

<!-- there is an extra backslash in script to show the original expression in bullet points 2,4,and 5 because the first backlash escapes -->

*   (.)\1\1
*   "(.)(.)\\\2\\\1".              
*   (..)\1
*   "(.).\\\1.\\\1"               
*   "(.)(.)(.).*\\\3\\2\\\1"         


<font size=3><b> Solution to Question 3 </font></b>

<b><i> there is an extra backslash in script to show the original expression in bullet points 2,4,and 5 because the first backlash escapes ; in question above is original expression. </b></i>


*  (.)\1\1 -  matches any character that is repeated three times in a row, like ‘aaa’, 'bbb'.

*  "(.)(.)\\\2\\\1" - matches any two characters followed by the same two characters in reverse order, example: abba or 1441.

*  (..)\1 - this expression will match with 2 adjoining repeats of a pair of characters or two characters repeated; example '3131' or 'abab' . 

*  “(.).\\\1.\\\1” - this match, will search for a five character term, three of which are the same. Example lets say you have characters 'd' 'e' 'f'. The output would be  'dedfd' or '41424'.

*   “(.)(.)(.).*\\\3\\\2\\\1” - this will match a set of characters that begin and end with the same three characters, except the second instance is reversed, like “racecar” or '12321'.



## Question 4 

Construct regular expressions to match words that:

* Start and end with the same character.
* Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
* Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

<font size=3><b> Solution to Question 4 </font></b>


<b><i> there is an extra backslash in script to show the expression I intend to be solution. a to c is  2 backslash but made it 3 backlashes in script to show 2 backlashes for html </b></i>

<!-- there is an extra backslash in script to show the original expression; read above -->

a).  Start and end with the same character ---> ^(.).*\\\1$ 

b). Contain a repeated pair of letters ---->  (.).*\\\1 

c). Contain one letter repeated in at least three places ---> (.).*\\\1.*\\\1 

