---
title: "Week 2 lab 607"
author: "Blessing Anoroh"
date: "February 2, 2024"
output: html_document
---

## Week 2 Assignment R and SQL 

Part 1 : Build a table

* Choose six recent popular movies. 
* Ask at least five people that you know (friends, family, classmates, imaginary friends if necessary) to rate each of these movies that they have seen on a scale of 1 to 5.


Part 2: Store data in SQL database 

* Take the results (observations) and store them in the class MySQL database: \
-Server name: cunydata607sql.mysql.database.azure.com  \
-Username / password: will be given to you in an email \
Note: it is good practice to change your password. To do so, use this SQL command: 	SET PASSWORD = ‘<your new password here>';

Part 3: Transfer data from SQL database to R dataframe 

* Load the information from the SQL database into an R dataframe.


<font size =3> Build Table and Store Data in SQL Database </font>

A table has been created and stored in the cunydata607sql.mysql.database.azure.com called “movieratings”.



```{r setup}
library (tidyverse)
library(openintro)
library (RMySQL)
library(DBI)
```



```{r}
#Import from SQL reference to Part 3

mysqldb = dbConnect(RMySQL::MySQL(),
                            dbname= 'blessing.abraham-anoroh65',
                            host= 'cunydata607sql.mysql.database.azure.com',
                            port=3306,
                            user='blessing.abraham-anoroh65',
                            password='blessing.abraham-anoroh65')

dbListTables(mysqldb)
```



```{r, results}
#fetching results from SQL

result = dbSendQuery(mysqldb, "select * from movieratings")

movieratings <- fetch(result)

print(movieratings)
```


Part 4: Missing data strategy 

* Implement an approach to missing data 
* Explain why you decided to take the chosen approach \
Note: consider that later in the course you will revisit this information you have collected and will use it to implement a Recommend-er.

For missing data I would identify where the missing data is using WHERE command to find "NULL" - which is considered to empty perhaps. I can also remove or delete where there is missing data. But in this exercise I would input the missing data using the number "0" as the NULL because the rating is from 1 to 5. And I would not want the missing data to cause an issue in running the SQL.



