---
title: "Week 1 Lab 607"
author: "Blessing A"
output: html_document
date: "February 1, 2024"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library (readr);
library (tidyverse);
library (knitr);
library (dplyr);
library(ggplot2)
congressage <- read.csv(url("https://raw.githubusercontent.com/fivethirtyeight/data/master/congress-demographics/data_aging_congress.csv"))
```

<font> Overview: The article I chose is "Congress Today Than It's Ever Been". I choose this article because it ccaught my interest. I've been talking about this lately outside school; that many people in the office are just becoming older and older. And these older representative lack modern familiriaty with modern technology as the article stated. In this R I will show the trend in median age over time of representatives using graphs, and or plots. Link: https://fivethirtyeight.com/features/aging-congress-boomers/.</font>

<font size ="4"> Plot Attempt </font>

```{r congressage plotted}
ggplot(data = congressage, aes(x = congress, y = age_years)) +
  geom_jitter(width = 0.1, height = 0, size = 2, alpha = 0.6, color = "darkblue") +
  xlab("Congress") +
  ylab("Age") +
  geom_line()

```

<font size ="4"> Sorted Group and Plotting </font> 

```{r}
  
# Group by congress and calculated mean and median age      
  sortedCongress <- congressage %>%
  group_by(congress) %>%
  summarize(avg_age = mean(age_years, na.rm = TRUE),
            median_age = median(age_years, na.rm = TRUE))
  
# Plotting
  ggplot(sortedCongress, aes(x = congress)) +
  geom_point(aes(y = avg_age, color = "Average Age")) +
  geom_point(aes(y = median_age, color = "Median Age")) +
  xlab("Congress") +
  ylab("Age") +
  scale_color_manual(values = c("red", "blue"), guide = guide_legend(title = NULL)) +
  theme_minimal()
```

<font size ="4"> Line Graph for average and median age </font> 

```{r}
# Create a line graph for average and median age
ggplot(sortedCongress, aes(x = congress)) +
  geom_line(aes(y = avg_age, color = "Average Age")) +
  geom_line(aes(y = median_age, color = "Median Age")) +
  xlab("Congress") +
  ylab("Age") +
  scale_color_manual(values = c("red", "blue"), guide = guide_legend(title = NULL)) +
  theme_minimal()
```


<font size ="4"> Summary of Sorted Congress </font> 
```{r}
# Display the summary of sortedCongress
glimpse(sortedCongress)
```


<font size= "4"> Conclusion: Based on the plots, as congress became more and more(and th years increas) so did the age of representatives in Congress rise.</font>
