---
title: 'Week 5 607 : Unitdying Data'
author: "Blessing Anoroh"
date: "February 20, 2024"
output: html_document
---
\

(1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above.  You’re encouraged to use a “wide” structure similar to how the information appears above, so that you can practice tidying and transformations as described below. \
(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. \
(3) Perform analysis to compare the arrival delays for the two airlines. \
(4) Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data cleanup work, analysis, and conclusions.  Please include in your homework submission: \
* The URL to the .Rmd file in your GitHub repository. and  \
* The URL for your rpubs.com web page. 

```{r setup}
library(tidyr)
library(dplyr)
library(tidyverse)
```

<font size=5><b>  Creating a .CSV file and Read data </font></b>

```{r}
#create CSV
csv <- rbind(c(NA,NA , "Los Angeles", "Phoenix", "San Diego", "San Francisco", "Seattle"),
             c("ALASKA", "On Time", 497, 221, 212, 503, 1841),
             c(NA, "Delayed", 62, 12, 20, 102, 305),
             c("AM WEST", "On Time", 694, 4840, 383, 320, 201),
             c(NA, "Delayed", 117, 415, 65, 129, 61))

write.table(csv, file = "Wk3Assignment_Data_607.csv", sep = ",", col.names=F, row.names=F)

#read data using dplyr and tidyr
flight_data <- read.csv(paste0("Wk3Assignment_Data_607.csv"), 
                             stringsAsFactors = F); flight_data

```
\ 

<font size=5><b>  Tidying Data </font></b>

```{r}

#tidy by Airline and Status (delay or on time) with use of tidyr and dplyr
long_data2 <- flight_data %>%
mutate(Airlines1 = NA., Airlines2=lag(NA.)) %>% 
mutate(Airline = coalesce(Airlines1,Airlines2), Status = NA..1)  %>%
gather("Destination", "Flights", 3:7) %>% 
select(Airline:Flights) %>% 
arrange(Airline, desc(Status), Destination); long_data2
```

```{r}
#tidy by cities
flightdata2 <- flight_data %>% gather("city","no.flights", 3:7)
flightdata2
```
\
 
<font size=5><b> Performing analysis to compare the arrival delays for the two airlines </font></b>

```{r}
Perform <- long_data2 %>%
  group_by(Airline) %>%
  mutate(A_Total = sum(Flights)) %>% 
  group_by(Airline, Status) %>%
  mutate(S_Total = sum(Flights), A_Perf = S_Total / A_Total) %>% 
  group_by(Airline, Destination) %>%
  mutate(D_Total = sum(Flights), D_Perf = Flights / D_Total)

# both airlines AL and AM Overall Performance
data.frame(Perform[c(1,10,11,20), c(1,2,7)])
```

```{r}
# Alaska Airlines and AM West by Destination
Perf_by_Destination <- data.frame(Perform %>%
  filter(Status == "On Time") %>% 
  group_by(Destination) %>% 
  unite(temp, D_Total, D_Perf) %>%
  spread(Status, temp) %>%
  select(Destination, Airline, On_Time = `On Time`) %>% 
  spread(Airline, On_Time) %>%
  separate(ALASKA, into=c("Flights.AL", "Perf.AL"), sep = "_", convert=T) %>% 
  separate(`AM WEST`, into=c("Flights.AM", "Perf.AM"), sep = "_", convert=T) %>% 
  mutate(Difference = Perf.AL - Perf.AM)  %>% 
  arrange(Destination)); Perf_by_Destination
```