---
title: "Week 7 - Class 607"
author: "Blessing Anoroh"
date: "March 5, 2024"
output: html_document
---

## Assignment - Working with HTML, XML, and JSON

```{r setup}
library(XML)
library(htmltools)
library(jsonlite)
library(xml2)
library(DT)
library(htmlwidgets)
# for some R version DT may not be able to install  use the following below to install
# install.packages("htmlwidgets", type = "binary") and 
# install.packages("DT", type = "binary")
```
 
 \
<font size= 3> For each source I started with "read" directly from computer. But each Data are attached with source codes to have accessible urls. </font>
\

## Loading HTML Data into R: 

\
Source Code: https://github.com/baa5234/class607/blob/main/Week%207%20Assignment/mathbooks.html
\

```{r}
mathbooks_html <- readHTMLTable("/Users/blessinga/Desktop/Masters Data Science/week7_607/mathbooks.html")
mathbooks_html <- mathbooks_html[[1]]
datatable(mathbooks_html)
```


## Loading XML Data into R:

\
Source Code: https://github.com/baa5234/class607/blob/main/Week%207%20Assignment/mathbooks.json
\

```{r}
mathbooks_xml <- xmlToDataFrame("/Users/blessinga/Desktop/Masters Data Science/week7_607/mathbooks.xml")
datatable(mathbooks_xml)
```

## Loading JSON Data into R:

\
Source Code: https://github.com/baa5234/class607/blob/main/Week%207%20Assignment/mathbooks.xml

\

```{r}
 mathbooks_json_df <- "/Users/blessinga/Desktop/Masters Data Science/week7_607/mathbooks.json"
mathbooks_json2 <- jsonlite::fromJSON(mathbooks_json_df)
mathbooks_json <- as.data.frame(mathbooks_json2$books)
datatable(mathbooks_json)
```

## Conclusion: 

  Are the three data frames identical? 
  
  <font size = 4> <b> Yes the 3 data frames for HTML, XML, and Json are all identical. And also it can be tested using identical function. </b></font>

\
\
